Hydrodynamics and rheology of an adhesive vesicle doublet

### 2019-02-08 vesicle_code/docs/adhesion/adhesion.tex
edits from yny on Feb8

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Effects of the adhesion parameters on draining times}
\label{subsec:qflow_draining_times}
When two vesicles move towards (or away from) each other under a
constant force $F$ without any external fluid flow, the height, $h$, of
the thin liquid film between two vesicles follows the draining
dynamics~\cite{RamachandranLeal2010_PoF}
\begin{align}
  \label{eq:dhdt}
  \frac{d h}{dt} \sim \frac{K_a^{2/3} F^{1/3}}{\mu R_0^{10/3}} h^3,
\end{align}
where $K_a$ is the area expansion modulus, $\mu$ is the viscosity of the
exterior fluid, and $R_0$ is the radius of the undeformed vesicle.  For
the case of a constant forcing $F$ (independent of separation distance
$h$ and time $t$), equation~\eqref{eq:dhdt} can be easily integrated to
relate the film thickness $h$ and time $t$:
\begin{align*}
  t \sim \frac{\mu R_0^{10/3}}{K_a^{2/3} F^{1/3}}
    \frac{1}{2 h^2} \bigg|^{h(t)}_{h(0)},
\end{align*}
where $h(0)$ is the vesicle separation at $t=0$.  We note that $F< (>)
0$ for an attractive  (repulsive) interaction between two vesicles,  and
consequently $h(t)$ decreases (increases) from  the initial separation
distance $h(0)$.  When the force is attractive, $h$ decreases
monotonically, and $t$ in the above equation is the draining time that
diverges as $h\rightarrow 0$.

@@

When $\theta=\pi$ (the doublet long axis is aligned with the far-field diverging flow), the converging flow
pushes vesicles towards each other and the short-range repulsive force is essential to keep the two vesicles at the separation distance.
With the converging flow pushing the two vesicles towards the stagnation point dynamically placed at the doublet center, 
this configuration is expected to be more stable than the $\theta=\pi/2$ configuration in the middle of Figure~\ref{fig:InclinationAngle},  where  
the doublet long axis is parallel to the converging direction. This is because
when $\theta=\pi/2$ the diverging flow is pulling the vesicles away from each other as shown in the illustration. 
With the inclination angle fixed at $\theta=\pi/2$,
the vesicle doublet stays intact only when the adhesive force is strong enough to dominate over the fluid drag in this configuration.
Thus it is possible that the diverging flow may overcome adhesion and
break up the doublet. It is also reasonable to expect a vesicle doublet to rotate from $\theta=\pi/2$ to $\theta=\pi$,
such that the fluid flow is pushing them together rather than pulling them apart.

In the following numerical experiments we place a vesicle doublet with $\theta=\pi/2$, so that
the diverging flow may be strong enough to pull vesicles away from the doublet.
%In this configuration the flow is pulling the vesicles away from each other, and 
%is expected to be less stable than the $\theta=\pi$ configuration where the fluid flow is constantly pushing vesicles towards each other. 
%
At low extensional
rates, we expect the $\theta=\pi/2$ doublet to stay intact at the fluid
trap stagnation point.  On the other hand, the vesicle doublet may
become unstable and eventually separate at higher extensional rates.
Thus we expect there to exist a critical extensional rate $\chi_c$ above which
the vesicle doublet cannot stay bound under a given adhesion potential.
Therefore, the dependence of the critical extensional rate $\chi_c$ on the
adhesion potential and the mechanical vesicle properties provides a
means to probe the physics of membrane adhesion.

@@

In Figures~\ref{fig:extensional1}--\ref{fig:extensional3},
snapshots of a fluid trap containing a vesicle doublet of reduced area
$\Delta A = 0.7$ are simulated at different extensional rates.  While
this doublet remains bound for all the
considered extensional rates, a doublet with a reduced area of $\Delta A
= 0.8$ is split with a critical extensional rate $\chi_c \sim 0.1$
(Figure~\ref{fig:extensional4}).  
%
\begin{figure}[htp]
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image05.png}
  \caption{\label{fig:extensional1} A doublet inside a fluid trap.  The
  reduced area is $\Delta A = 0.7$, the Hamaker constant is $\mathcal{H} =
  0.7$, the adhesion length scale is $\delta = 0.4$, and the extensional
  rate is $\chi = 0.02$.  The center of the vesicle, denoted with a
  black square, is a stagnation point.}
  \end{figure}
  %
  For extensional rates and reduced areas that result in a bound vesicle doublet, 
we observe that the doublet always rotate from $\theta=\pi/2$ configuration towards
the more stable $\theta=\pi$ configuration, as shown in Figures~\ref{fig:extensional1}--\ref{fig:extensional3}.
We also observe that in most cases with moderate extensional flow rates the doublet falls short of 
aligning their long axis with the diverging direction.

  \begin{figure}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image05.png}
  \caption{\label{fig:extensional2} A doublet inside a fluid trap.  The
  reduced area is $\Delta A = 0.7$, the Hamaker constant is $\mathcal{H}
  = 0.7$, the adhesion length scale is $\delta = 0.4$, and the
  extensional rate is $\chi = 0.07$.  The center of the vesicle, denoted
  with a black square, is a stagnation point.}
  \end{figure}

 \begin{figure}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image05.png}
  \caption{\label{fig:extensional3} Two vesicles in an extensional flow.
  The reduced area is $\Delta A = 0.7$, the Hamaker constant is $\mathcal{H} =
  0.7$, the adhesion length scale is $\delta = 0.4$, and the extensional
  rate is $\chi = 0.1$.  The center of the vesicle, denoted with a black
  square, is a stagnation point.}
  \end{figure}

In Figure~\ref{fig:extensionalInclinationAngle}, we plot the inclination
angle of the doublet as a function of time for four different
extensional rates (left) and the final inclination angle for all
doublets that reach an equilibrium state (middle).  We observe that not
only do larger smaller extensional rates result in smaller inclination
angles, but smaller reduced areas also result in smaller inclination
angles.  Moreover, the critical Hamaker constant that determines if the
doublet remains trapped or is broken decreases with an increasing
extensional rate.
\begin{figure}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image05.png}
  \caption{\label{fig:extensional4} A doublet broken by an extensional
  flow.  The reduced area is $\Delta A = 0.8$, the Hamaker constant is
  $\mathcal{H} = 0.7$, the adhesion length scale is $\delta = 0.4$, and
  the extensional rate is $\chi = 0.1$.  Initially, the doublet tilts to
  an inclination angle, and then the extensional rate overcomes the
  adhesive force and the doublet is broken.  The center of the vesicle,
  denoted with a black square, is a stagnation point.}
\end{figure}
We summarize the final inclination angle of the
doublet in the right plot of Figure~\ref{fig:extensionalInclinationAngle}.
The size of the round dots are scaled to the final inclination angle as
defined in the right plot.  At smaller extensional rates, the vesicles
come closer to aligning their long axis with the diverging direction.
When the doublet is broken at reduced area $\Delta A = 0.80$, the
extensional rate is sufficiently large to align the long axis of the
vesicle with the diverging direction and then the vesicles separate
(Figure~\ref{fig:extensional4}).  For very low extensional rates, the
time horizon is insufficient for the doublet to reach an equilibrium
state, and these simulations are marked with a square.
%
The right plot of Figure~\ref{fig:extensionalInclinationAngle} 
also summarizes the  reduced areas and
extensional rates that result in a bound vesicle doublet in a fluid trap: Parameter values with a blue
mark result in a fluid trap and parameter values with a red mark result
in vesicle separation.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Effects of the adhesion parameters on draining times}
\label{subsec:qflow_draining_times}
When two vesicles move towards (or away from) each other under a
constant force $F$ without any imposed external flow, the height, $h$, of
the thin liquid film between two vesicles follows the draining
dynamics~\cite{RamachandranLeal2010_PoF}
\begin{align}
  \label{eq:dhdt}
  \frac{d h}{dt} \sim \frac{K_a^{2/3} F^{1/3}}{\mu R_0^{10/3}} h^3,
\end{align}
where $K_a$ is the area expansion modulus, $\mu$ is the viscosity of the
exterior fluid, and $R_0$ is the radius of the undeformed vesicle.  For
the case of a constant forcing $F$ (independent of separation distance
$h$ and time $t$), equation~\eqref{eq:dhdt} can be easily integrated to
relate the film thickness $h$ and time $t$:
\begin{align*}
  t \sim \frac{\mu R_0^{10/3}}{K_a^{2/3} F^{1/3}}
    \frac{1}{2 h^2} \bigg|^{h(t)}_{h(0)},
\end{align*}
where $h(0)$ is the vesicle separation at $t=0$.  We note that $F< (>)
0$ for an attractive  (repulsive) interaction between two vesicles,  and
consequently $h(t)$ decreases (increases) from  the initial separation
distance $h(0)$.  When the force is attractive, $h$ decreases
monotonically, and $t$ in the above equation is the draining time that
diverges as $h\rightarrow 0$.

@@

When $\theta=\pi$ (the doublet long axis is aligned with the far-field diverging flow), the converging flow
pushes vesicles towards each other and the short-range repulsive force is essential to keep the two vesicles at the separation distance.
With the converging flow pushing the two vesicles towards the stagnation point dynamically placed at the doublet center, 
this configuration is expected to be more stable than the $\theta=\pi/2$ configuration in the middle of Figure~\ref{fig:InclinationAngle},  where  
the doublet long axis is parallel to the converging direction. This is because
when $\theta=\pi/2$ the diverging flow is pulling the vesicles away from each other as shown in the illustration. 
If the doublet inclination angle is fixed at $\theta=\pi/2$,
the vesicle doublet stays intact only when the adhesive force is strong enough to dominate over the fluid drag in this configuration.
Thus it is possible that the diverging flow may overcome adhesion and
break up the doublet. It is also reasonable to expect a vesicle doublet to rotate from $\theta=\pi/2$ to $\theta=\pi$,
such that the fluid flow is pushing them together rather than pulling them apart.

In the following numerical experiments we place a vesicle doublet with $\theta=\pi/2$, so that
the diverging flow may be strong enough to pull vesicles away from the doublet.
%In this configuration the flow is pulling the vesicles away from each other, and 
%is expected to be less stable than the $\theta=\pi$ configuration where the fluid flow is constantly pushing vesicles towards each other. 
%
At low extensional
rates, we expect the $\theta=\pi/2$ doublet to stay bound at the fluid
trap stagnation point.  On the other hand, the vesicle doublet may
become unstable and eventually separate at higher extensional rates.
Thus we expect there to exist a critical extensional rate $\chi_c$ above which
the vesicle doublet cannot stay bound under a given adhesion potential.
Therefore, the dependence of the critical extensional rate $\chi_c$ on the
adhesion potential and the mechanical vesicle properties provides a
means to probe the physics of membrane adhesion.

@@

In Figures~\ref{fig:extensional1},
snapshots of a fluid trap containing a vesicle doublet of reduced area
$\Delta A = 0.7$ are simulated at different extensional rates.  While
this doublet remains bound for all the
considered extensional rates, a doublet with a reduced area of $\Delta A
= 0.8$ is split with a critical extensional rate $\chi_c =0.1$
(bottom row in Figure~\ref{fig:extensional3}).  
%
\begin{figure}[htp]
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image05.png}\\
   \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image05.png}
  \caption{\label{fig:extensional1} A doublet inside a fluid trap.  The
  reduced area is $\Delta A = 0.7$, the Hamaker constant is $\mathcal{H} =
  0.7$, the adhesion length scale is $\delta = 0.4$, and the extensional
  rate $\chi = 0.02$ ({\em top}) and $\chi=0.7$ ({\em bottom}).  The center of the vesicle, denoted with a
  black square, is a stagnation point.}
  \end{figure}
  %
  For extensional rates and reduced areas that result in a bound vesicle doublet, 
we observe that the doublet always rotate from $\theta=\pi/2$ configuration towards
the more stable $\theta=\pi$ configuration, as shown in Figures~\ref{fig:extensional1}--\ref{fig:extensional3}.
We also observe that in most cases with moderate extensional flow rates the doublet falls short of 
aligning their long axis with the diverging direction.

%  \begin{figure}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image01.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image02.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image03.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image04.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image05.png}
%  \caption{\label{fig:extensional2} A doublet inside a fluid trap.  The
%  reduced area is $\Delta A = 0.7$, the Hamaker constant is $\mathcal{H}
%  = 0.7$, the adhesion length scale is $\delta = 0.4$, and the
%  extensional rate is $\chi = 0.07$.  The center of the vesicle, denoted
%  with a black square, is a stagnation point.}
%  \end{figure}

 \begin{figure}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image05.png}\\
    \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image05.png}
  \caption{\label{fig:extensional3} Two vesicles in an extensional flow.
  The reduced area $\Delta A = 0.7$ ({\em top}) and $\Delta A=0.8$ ({\em bottom}). The Hamaker constant is $\mathcal{H} =
  0.7$, the adhesion length scale is $\delta = 0.4$, and the extensional
  rate $\chi = 0.1$.  The center of the vesicle, denoted with a black
  square, is a stagnation point.}
  \end{figure}

In Figure~\ref{fig:extensionalInclinationAngle}, we plot the inclination
angle of the doublet as a function of time for four different
extensional rates (left) and the final inclination angle for all
doublets that reach an equilibrium state (middle).  We observe that not
only do larger smaller extensional rates result in smaller inclination
angles, but smaller reduced areas also result in smaller inclination
angles.  Moreover, the critical Hamaker constant that determines if the
doublet remains trapped or is broken decreases with an increasing
extensional rate.
%\begin{figure}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image01.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image02.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image03.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image04.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image05.png}
%  \caption{\label{fig:extensional4} A doublet broken by an extensional
%  flow.  The reduced area is $\Delta A = 0.8$, the Hamaker constant is
%  $\mathcal{H} = 0.7$, the adhesion length scale is $\delta = 0.4$, and
%  the extensional rate is $\chi = 0.1$.  Initially, the doublet tilts to
%  an inclination angle, and then the extensional rate overcomes the
%  adhesive force and the doublet is broken.  The center of the vesicle,
%  denoted with a black square, is a stagnation point.}
%\end{figure}
We summarize the final inclination angle of the
doublet in the right plot of Figure~\ref{fig:extensionalInclinationAngle}.
The size of the round dots are scaled to the final inclination angle as
defined in the right plot.  At smaller extensional rates, the vesicles
come closer to aligning their long axis with the diverging direction.
When the doublet is broken at reduced area $\Delta A = 0.80$, the
extensional rate is sufficiently large to align the long axis of the
vesicle with the diverging direction and then the vesicles separate
(bottom row in Figure~\ref{fig:extensional3}).  For very low extensional rates, the
time horizon is insufficient for the doublet to reach an equilibrium
state, and these simulations are marked with a square.
%
The right plot of Figure~\ref{fig:extensionalInclinationAngle} 
also summarizes the  reduced areas and
extensional rates that result in a bound vesicle doublet in a fluid trap: Parameter values with a blue
mark result in a fluid trap and parameter values with a red mark result
in vesicle separation.
